Meta-analysis

### 2019-09-14 05_MetaAnalysis.Rmd
Fix date formats on import

# convert coltypes

PM_MB_means <- 
   PM_MB_means %>% 
   mutate(year = as.integer(year)) %>% 
   mutate(replicates = as.integer(replicates)) %>% 
   mutate(planting_date = as.Date(planting_date)) %>%
   mutate(emergence_date = as.Date(emergence_date)) %>% 
   mutate(flowering_date = as.Date(flowering_date)) %>% 
   mutate(pod_final_date = as.Date(pod_final_date)) %>% 
   mutate(mid_late_pod_final = as.Date(mid_late_pod_final)) %>% 
   mutate(first_sign_disease = as.Date(first_sign_disease)) %>% 
   mutate(fungicide_application_1 = as.Date(fungicide_application_1)) %>% 
   mutate(fungicide_application_2 = as.Date(fungicide_application_2)) %>% 
   mutate(fungicide_application_3 = as.Date(fungicide_application_3)) %>%
   mutate(fungicide_application_4 = as.Date(fungicide_application_4)) %>% 
   mutate(fungicide_application_5 = as.Date(fungicide_application_5)) %>%
   mutate(fungicide_application_6 = as.Date(fungicide_application_6)) %>%
   mutate(fungicide_application_7 = as.Date(fungicide_application_7)) %>% 
   mutate(harvest_date = as.Date(harvest_date)) %>% 
   mutate(final_assessment = as.Date(final_assessment))

@@

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Replicates) # degrees of freedom within groups
   # n.A <- PM_MB_means[PM_MB_means$trial_ref == "Trial2012,]$Replicates[1]
   # n.B <- PM_MB_means[PM_MB_means$trial_ref == "Trial2012,]$Replicates[2]
   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Yield_error[1]

@@

   #  LSD = (Tcrit) * sqrt(MSW * (1/n.A + 1/n.B))
   #  (LSD/Tcrit) = sqrt(MSW * (1/n.A + 1/n.B))
   #  (LSD/Tcrit)^2 = MSW * (1/n.A + 1/n.B)
   #  (LSD/Tcrit)^2/(1/n.A + 1/n.B) = MSW
   #  MSE <- (LSD/Tcrit)^2/(1/n.A + 1/n.B) # This formula was abandon in favor for the formula in Nugugi et. at (2011)
   MSE <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Replicates[1] * (LSD /
                                                                           Tcrit) ^ 2 / 2 # method for finding variance in Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Replicates))
      }
      #  Sv[j]^2 = MSE/(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Replicates[j] *
      #                   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$grain_yield.t.ha.[j])
      # Balance equation
      Sv[j] = sqrt(MSE / (
         PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Replicates[j] *
            as.numeric(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$grain_yield.t.ha.[j])
      ))

@@

   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$y_error.type <-
      "SamplingVariance"

@@

for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$Yield_error)) {
   # create empty vector at first iteration with length of j
   if (j == 1) {
      Sv <-
         vector(length = length(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$Yield_error))
   }

   # Equation to convert standard deviation to Sampling Variance
   Sv[j] = (PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$Yield_error[j]) ^
      2
}

# and replacing the values in the dataset
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$Yield_error <-
   Sv
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$y_error.type <-
   "SamplingVariance"

@@

# for(i in PM_MB_means$trial_ref){
#
# log(V)[i] = a + b * log(
#    mean(subset(PM_MB_means$grain_yield.t.ha.[i], fungicide_ai == "control"),
#         subset(PM_MB_means$grain_yield.t.ha.[i], fungicide_ai == "tebuconazole")))
# # need to know what parameters a and b are here
# # I am guessing they represent prediction parameters for each "study" for each of the treatments, "control" and tebuconazole"
# }

@@

plot(PM_MB_train$Yield_error,
     log(PM_MB_train$`grain_yield.t.ha.`))

# Run a linear model with Grain_yield as a fixed effect predictor and Trial as a random intercept
PM_MB_trianLMER <- lmer(Yield_error ~ `grain_yield.t.ha.` +
                           (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trianLMER)

@@

PM_MB_trianLMER.1 <- lmer(log(Yield_error) ~ `grain_yield.t.ha.` +
                             (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trianLMER.1) # grain yeild a significant predictor of log(yeild error)
anova(PM_MB_trianLMER, PM_MB_trianLMER.1) # no significant difference between the models

PM_MB_trianLMER.2 <- lmer(Yield_error ~ log(`grain_yield.t.ha.`) +
                             (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trianLMER.2) # log(grain yeild) is a significant predictor of yeild error
anova(PM_MB_trianLMER, PM_MB_trianLMER.2) # no significant difference between the models

PM_MB_trianLMER.3 <-
   lmer(log(Yield_error) ~ log(`grain_yield.t.ha.`) +
           (1 | trial_ref),
        data = PM_MB_train)
summary(PM_MB_trianLMER.3) # log(grain yeild) is a significant predictor of log(yeild error)
anova(PM_MB_trianLMER, PM_MB_trianLMER.3) # no significant difference between the models

# Lets try with Fungicide as a fixed effect
PM_MB_trianLMER.4 <-
   lmer(Yield_error ~ `grain_yield.t.ha.` + as.factor(Fungicide) +
           (1 | trial_ref),
        data = PM_MB_train)
summary(PM_MB_trianLMER.4)
anova(PM_MB_trianLMER, PM_MB_trianLMER.4) # no significant difference between the models

@@

# plot of the mixed effect model, Black line shows the mean lmer regression line
ggplot(PM_MB_train) +
   geom_point(aes(y = Yield_error, x = `grain_yield.t.ha.`, colour = trial_ref)) +
   geom_abline(
      intercept = c(InterC, lmer.coef$trial_ref[, 1]),
      slope = c(paramA, lmer.coef$trial_ref[, 2]),
      size = 1,
      colour = c("Black", hue_pal()(11))
   ) +
   coord_cartesian(xlim = c(0, 3))

@@

for (i in NATrials$trial_ref) {
   #Sv <- (log(PM_MB_means[PM_MB_means$trial_ref == i,]$`grain_yield.t.ha.`) - InterC)/ paramA  # parameter from when I had the predictor and response variable flipped

   Sv <-
      (PM_MB_means[PM_MB_means$trial_ref == i,]$`grain_yield.t.ha.` * paramA) + InterC

   PM_MB_means[PM_MB_means$trial_ref == i,]$Yield_error <- Sv
   PM_MB_means[PM_MB_means$trial_ref == i,]$y_error.type <-
      "SamplingVariance"
}

@@

# Rerun the model
PM_MB_trianLMER.5 <- lmer(Yield_error ~ `grain_yield.t.ha.` +
                             (1 | trial_ref), data = PM_MB_train.2)
summary(PM_MB_trianLMER.5) # no significant effect of mean grain yield on yeild error

@@

# plot the data with the predicted slopes overlayed
ggplot(PM_MB_train.2) +
   geom_point(aes(y = Yield_error, x = `grain_yield.t.ha.`, colour = trial_ref)) +
   geom_abline(
      intercept = c(InterC, lmer.coef$trial_ref[, 1]),
      slope = c(paramA, lmer.coef$trial_ref[, 2]),
      size = 1,
      colour = c("Black", hue_pal()(10))
   ) +
   coord_cartesian(ylim = c(0, 0.33))

@@

Just for kicks lets look at the linear regression model as a Random effects model where `grain_yield.t.ha.` is set as a random slope to the random intercept `trial_ref`.  
The model fits better as shown by an ANOVA comparison. But how to impute from this model if mean variance for a trial is unknown....
```{r random_slope_model}

# Run a linear model with Yield error as a fixed effect predictor and Trial as a random intercept
PM_MB_trianLMERx <- lmer(Yield_error ~ 1 +
                            (`grain_yield.t.ha.` | trial_ref), data = PM_MB_train)

@@

# plot of the mixed effect model, Black line shows the mean lmer regression line
ggplot(PM_MB_train) +
   geom_point(aes(y = Yield_error, x = `grain_yield.t.ha.`, colour = trial_ref)) +
   geom_abline(
      intercept = lmer.coefx$trial_ref[, 2],
      slope = lmer.coefx$trial_ref[, 1],
      size = 1,
      colour = hue_pal()(11)
   ) +
   coord_cartesian(xlim = c(0, 3),
                   ylim = c(0, 1))
```

@@

meta_mean.GY <- mean(PM_MB_means.SV$grain_yield.t.ha., na.rm = TRUE)

PM_MB_means_T <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(
      trial_mean = mean(grain_yield.t.ha., na.rm = TRUE),
      BW_trial_variance = sd(grain_yield.t.ha., na.rm = TRUE) ^ 2
   )
PM_MB_means_T

>>>>

# convert coltypes
PM_MB_means <-
   PM_MB_means %>%
   mutate(year = as.integer(year)) %>%
   mutate(replicates = as.integer(replicates)) %>%
   mutate(planting_date = as.Date(planting_date, format = "%d/%m/%Y")) %>%
   mutate(emergence_date = as.Date(emergence_date, format = "%d/%m/%Y")) %>%
   mutate(flowering_date = as.Date(flowering_date, format = "%d/%m/%Y")) %>%
   mutate(pod_final_date = as.Date(pod_final_date, format = "%d/%m/%Y")) %>%
   mutate(mid_late_pod_final = as.Date(mid_late_pod_final, format = "%d/%m/%Y")) %>%
   mutate(first_sign_disease = as.Date(first_sign_disease, format = "%d/%m/%Y")) %>%
   mutate(fungicide_application_1 = as.Date(fungicide_application_1, format = "%d/%m/%Y")) %>%
   mutate(fungicide_application_2 = as.Date(fungicide_application_2, format = "%d/%m/%Y")) %>%
   mutate(fungicide_application_3 = as.Date(fungicide_application_3, format = "%d/%m/%Y")) %>%
   mutate(fungicide_application_4 = as.Date(fungicide_application_4, format = "%d/%m/%Y")) %>%
   mutate(fungicide_application_5 = as.Date(fungicide_application_5, format = "%d/%m/%Y")) %>%
   mutate(fungicide_application_6 = as.Date(fungicide_application_6, format = "%d/%m/%Y")) %>%
   mutate(fungicide_application_7 = as.Date(fungicide_application_7, format = "%d/%m/%Y")) %>%
   mutate(harvest_date = as.Date(harvest_date, format = "%d/%m/%Y")) %>%
   mutate(final_assessment = as.Date(final_assessment, format = "%d/%m/%Y"))
```

@@

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) # degrees of freedom within groups
   # n.A <- PM_MB_means[PM_MB_means$trial_ref == "Trial2012,]$replicates[1]
   # n.B <- PM_MB_means[PM_MB_means$trial_ref == "Trial2012,]$replicates[2]
   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error[1]

@@

   #  LSD = (Tcrit) * sqrt(MSW * (1/n.A + 1/n.B))
   #  (LSD/Tcrit) = sqrt(MSW * (1/n.A + 1/n.B))
   #  (LSD/Tcrit)^2 = MSW * (1/n.A + 1/n.B)
   #  (LSD/Tcrit)^2/(1/n.A + 1/n.B) = MSW
   #  MSE <- (LSD/Tcrit)^2/(1/n.A + 1/n.B) # This formula was abandon in favor for the formula in Nugugi et. at (2011)
   MSE <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[1] * (LSD /
                                                                            Tcrit) ^ 2 / 2 # method for finding variance in Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates))
      }
      #  Sv[j]^2 = MSE/(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates[j] *
      #                   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$grain_yield.t.ha[j])
      # Balance equation
      Sv[j] = sqrt(MSE / (
         PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[j] *
            as.numeric(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$grain_yield.t.ha[j])
      ))

@@

   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$y_error.type <-
      "SamplingVariance"

@@

for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error)) {
   # create empty vector at first iteration with length of j
   if (j == 1) {
      Sv <-
         vector(length = length(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error))
   }

   # Equation to convert standard deviation to Sampling Variance
   Sv[j] = (PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error[j]) ^
      2
}

# and replacing the values in the dataset
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error <-
   Sv
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$y_error.type <-
   "SamplingVariance"

@@

# for(i in PM_MB_means$trial_ref){
#
# log(V)[i] = a + b * log(
#    mean(subset(PM_MB_means$grain_yield.t.ha[i], fungicide_ai == "control"),
#         subset(PM_MB_means$grain_yield.t.ha[i], fungicide_ai == "tebuconazole")))
# # need to know what parameters a and b are here
# # I am guessing they represent prediction parameters for each "study" for each of the treatments, "control" and tebuconazole"
# }

@@

plot(PM_MB_train$yield_error,
     log(PM_MB_train$grain_yield.t.ha))

# Run a linear model with grain_yield.t.ha as a fixed effect predictor and trial as a random intercept
PM_MB_trianLMER <- lmer(yield_error ~ grain_yield.t.ha +
                           (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trianLMER)

@@

PM_MB_trianLMER.1 <- lmer(log(yield_error) ~ `grain_yield.t.ha` +
                             (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trianLMER.1) # grain yeild a significant predictor of log(yeild error)
anova(PM_MB_trianLMER, PM_MB_trianLMER.1) # no significant difference between the models

PM_MB_trianLMER.2 <- lmer(yield_error ~ log(`grain_yield.t.ha`) +
                             (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trianLMER.2) # log(grain yeild) is a significant predictor of yeild error
anova(PM_MB_trianLMER, PM_MB_trianLMER.2) # no significant difference between the models

PM_MB_trianLMER.3 <-
   lmer(log(yield_error) ~ log(`grain_yield.t.ha`) +
           (1 | trial_ref),
        data = PM_MB_train)
summary(PM_MB_trianLMER.3) # log(grain yeild) is a significant predictor of log(yeild error)
anova(PM_MB_trianLMER, PM_MB_trianLMER.3) # no significant difference between the models

# Lets try with Fungicide as a fixed effect
PM_MB_trianLMER.4 <-
   lmer(yield_error ~ `grain_yield.t.ha` + as.factor(fungicide_ai) +
           (1 | trial_ref),
        data = PM_MB_train)
summary(PM_MB_trianLMER.4)
anova(PM_MB_trianLMER, PM_MB_trianLMER.4) # no significant difference between the models

@@

# plot of the mixed effect model, Black line shows the mean lmer regression line
ggplot(PM_MB_train) +
   geom_point(aes(y = yield_error, x = `grain_yield.t.ha`, colour = trial_ref)) +
   geom_abline(
      intercept = c(InterC, lmer.coef$trial_ref[, 1]),
      slope = c(paramA, lmer.coef$trial_ref[, 2]),
      size = 1,
      colour = c("Black", hue_pal()(11))
   ) +
   coord_cartesian(xlim = c(0, 3))

@@

for (i in NATrials$trial_ref) {
   #Sv <- (log(PM_MB_means[PM_MB_means$trial_ref == i,]$`grain_yield.t.ha`) - InterC)/ paramA  # parameter from when I had the predictor and response variable flipped

   Sv <-
      (PM_MB_means[PM_MB_means$trial_ref == i,]$`grain_yield.t.ha` * paramA) + InterC

   PM_MB_means[PM_MB_means$trial_ref == i,]$yield_error <- Sv
   PM_MB_means[PM_MB_means$trial_ref == i,]$y_error.type <-
      "SamplingVariance"
}

@@

# Rerun the model
PM_MB_trianLMER.5 <- lmer(yield_error ~ `grain_yield.t.ha` +
                             (1 | trial_ref), data = PM_MB_train.2)
summary(PM_MB_trianLMER.5) # no significant effect of mean grain yield on yeild error

@@

# plot the data with the predicted slopes overlayed
ggplot(PM_MB_train.2) +
   geom_point(aes(y = yield_error, x = `grain_yield.t.ha`, colour = trial_ref)) +
   geom_abline(
      intercept = c(InterC, lmer.coef$trial_ref[, 1]),
      slope = c(paramA, lmer.coef$trial_ref[, 2]),
      size = 1,
      colour = c("Black", hue_pal()(10))
   ) +
   coord_cartesian(ylim = c(0, 0.33))

@@

Just for kicks lets look at the linear regression model as a Random effects model where `grain_yield.t.ha` is set as a random slope to the random intercept `trial_ref`.  
The model fits better as shown by an ANOVA comparison. But how to impute from this model if mean variance for a trial is unknown....
```{r random_slope_model}

# Run a linear model with Yield error as a fixed effect predictor and Trial as a random intercept
PM_MB_trianLMERx <- lmer(yield_error ~ 1 +
                            (`grain_yield.t.ha` | trial_ref), data = PM_MB_train)

@@

# plot of the mixed effect model, Black line shows the mean lmer regression line
ggplot(PM_MB_train) +
   geom_point(aes(y = yield_error, x = `grain_yield.t.ha`, colour = trial_ref)) +
   geom_abline(
      intercept = lmer.coefx$trial_ref[, 2],
      slope = lmer.coefx$trial_ref[, 1],
      size = 1,
      colour = hue_pal()(11)
   ) +
   coord_cartesian(xlim = c(0, 3),
                   ylim = c(0, 1))
```

@@

meta_mean.GY <- mean(PM_MB_means.SV$grain_yield.t.ha, na.rm = TRUE)

PM_MB_means_T <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(
      trial_mean = mean(grain_yield.t.ha, na.rm = TRUE),
      BW_trial_variance = sd(grain_yield.t.ha, na.rm = TRUE) ^ 2
   )
PM_MB_means_T
